Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-10-14 paper/main.tex
notes

We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}. Sheaves are mathematical abstractions of how data over different parts of a continuous base space are glued together \cite{ghristElementaryAppliedTopology2014, ghri}, whether the data is distributed, streaming, or in disk; therefore sheaves are a way to express the continuity that the artist preserves. 
We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema. The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist\ is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.

@@

We encode individual datasets in the fiberbundle as sections of the bundle. We follow the convention that $\Gamma(U_i, \dtotal)$ denotes the set of sections $\{\dsection^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over an open set $U_{i}\subseteq \dbase$.  As mentioned in \autoref{sec:related-work:fiber-bundles}, the fiber space is embedded in the total space $\dfiber \hookrightarrow \dtotal$. The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$ \cite{SheafMathematics2021,spanier1989algebraic}. Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\. For example, Spivak describes tables that have the same schema and discrete index as sections of the same fiber bundle\cite{spivakSIMPLICIALDATABASES}.

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\mathcal{\dtotal}}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\bullet: \dtotal\rightarrow \dtotal$, which are the fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ when the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$
  \end{itemize}
\end{definition}

The functor object of $\mathcal{E}$ is a presheaf because the definition of a presheaf is that it is a contravariant functor from a category $\mathcal{K}$ to a set \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. As illustrated in the diagram, 
\begin{equation}
  \label{eq:artist:presheaf}
  \begin{tikzcd}[row sep=small]
    U_1 \arrow[dd, "\iota"'] & {\Gamma(U_1, \dtotal)}                              \\
    {} \arrow[r, "\sheaf"]   & {}                                                  \\
    U_2                      & {\Gamma(U_2, \dtotal)} \arrow[uu, "\iota^*"']
    \end{tikzcd}
\end{equation}
the functor \sheaf\ is contravariant because the inclusion map $\iota$ between open sets in $\mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}. The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$. On the other hand, a set of functions defined on a space $U_2$ must be defined on the subspace $\iota: \Gamma(U_2, \dbase) \rightarrow \Gamma(U_1, \dbase)$ but the reverse does not need to be true. For example, a continuous function defined over the interval $\left[.25,.5\right]$ need not be defined over $\left[0, 1\right]$ but must exist over  $\left[.3, .45\right]$. We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf. We also denote the sheaf over the fiberbundle \dtotal\ as $\sheaf(\dtotal)$. A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}. The sheaf over a limit of open sets that contain a point \dbasepoint\ is approximately the same as a sheaf over a point \dbase\ and is called the stalk $\mathscr{F}_{k}\coloneqq \lim\limits_{U\ni k} \mathscr{F}(U)$ \cite{StalkSheaf2019}. The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$. Our criteria for equivariance is that the set of maps from fiber to fiber $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ is preserved, since these maps include the types of data transformations described in \autoref{sec:related-work:equivariance}. When a fiber bundle is trivial, meaning $\dtotal=\dbase \times \dfiber$, then we can dispense with localization and directly consider the fiber maps $Hom(F_{k}, F_{k})$. In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.

The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$, which we also denote as $\sheaf(\gtotal)$. The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase\, and $\Gamma(\mathcal{S}, \gtotal)$ is the set of sections $\gsection$ over $\mathcal{\gbase}$. The artist is a morphism of sheafs 
\begin{equation}
  \label{eq:artist:morphism-of-sheaf}
  \begin{tikzcd}
    & \Gamma(\dtotal) & & & \Gamma(\gtotal) & \\
{} \arrow[rr, "f", Rightarrow, shift right, shorten=3em] & 
& 
{} \arrow[r, "\vartist", shorten=-1.25em, bold] & {} \arrow[rr, "g", Rightarrow, shift right, shorten=3em] & & {} \\
& \mathcal{\dbase}^{op} \arrow[uu, "\mathcal{\dtotal}\coloneqq\sheaf_{\mathcal{\dtotal}}", bend left] 
  \arrow[uu, "\sheaf_{\mathcal{\dtotal}}"', bend right] 
  & & 
& \mathcal{\gbase}^{op} \arrow[uu, "\sheaf_{\mathcal{\gtotal}}", bend left] 
  \arrow[uu, "\mathcal{\gtotal}\coloneqq \sheaf_{\mathcal{\gtotal}}"', bend right] &   
  \end{tikzcd}
\end{equation}

therefore the artist is a map between functors
\begin{equation}
  \tag{\ref{eq:artist:artist}}
  \vartist:\mathcal{\dtotal}\rightarrow{\mathcal{\gtotal}} 
\end{equation} which means that the artist is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}. One of the properties of a natural transformation is that the following diagram commutes
\begin{equation}
  \label{eg:artist:natural-transformation}
  \begin{tikzcd}
    \mathcal{\dtotal} \arrow[d, "f"'] \arrow[r, "\vartist"] & \mathcal{\gtotal} \arrow[d, "\vartist_{f} \coloneqq g"] \\
    \mathcal{\dtotal} \arrow[r, "\vartist"]                 & \mathcal{\gtotal}                                      
  \end{tikzcd}
\end{equation}

meaning that functions $f: \mathcal{\dtotal}\rightarrow \mathcal{\dtotal}$

@@

such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}.

@@

\subsection{Union of Artists}
\label{sec:artist:union}
\begin{figure}[!h]
\centering
\subfloat[+]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
\label{fig:artist:union:glue}}
%\hfil
\subfloat[$\oplus$]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
\label{fig:artist:union:continuity}}
\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space \gtotal. The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image. \note{add a border/make the same size/etc so 1 doesn't run over into two}}
\end{figure}

@@

\section{Constraints: Construction and formal properties of Artists}
\label{sec:artist:construction}
\begin{figure*}[h!]
  \includegraphics[width=\linewidth]{q.png}
  \caption{}
  \label{fig:constraints:q-overall}
\end{figure*}

>>>>

\note{roadmap the rest of this section}

\subsection{Categorical Artist}
We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}. Sheaves are mathematical abstractions of how data over different parts of a continuous base space are glued together \cite{ghristElementaryAppliedTopology2014, ghri}, whether the data is distributed, streaming, or in disk; therefore sheaves are a way to express the continuity that the artist preserves. 
We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema. The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist\ is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.

@@

%%define a category of sections here
We encode individual datasets in the fiberbundle as sections of the bundle. We follow the convention that $\Gamma(U_i, \dtotal)$ denotes the set of sections $\{\dsection^{j}:U_i \rightarrow \dfiber_{U_{i}}\}_{j\in I}$ of the fiber bundle \dtotal\ over an open set $U_{i}\subseteq \dbase$.  As mentioned in \autoref{sec:related-work:fiber-bundles}, the fiber space is embedded in the total space $\dfiber \hookrightarrow \dtotal$. The sections $\Gamma(\dbase, \dtotal)$ are called the global sections of $\dtotal$ \cite{SheafMathematics2021,spanier1989algebraic}. Each section is how we represent a unique dataset, where the set of sections $\Gamma$ is the set of all datasets that have the same data fields in the \dfiber\ and continuity as encoded in\dbase\. For example, Spivak describes tables that have the same schema and discrete index as sections of the same fiber bundle\cite{spivakSIMPLICIALDATABASES}.

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf_{\mathcal{\dtotal}}: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
    \item \textit{morphisms} arbitrary endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $f: \dtotal\rightarrow \dtotal$, which are the fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ when the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$
  \end{itemize}
\end{definition}

The functor object of $\mathcal{E}$ is a presheaf because the definition of a presheaf is that it is a contravariant functor from a category to a category of sets \textbf{Set} \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. As illustrated in the diagram, 
\begin{equation}
  \label{eq:artist:presheaf}
  \begin{tikzcd}[row sep=small]
    U_1 \arrow[dd, "\iota"'] & {\Gamma(U_1, \dtotal)}                              \\
    {} \arrow[r, "\sheaf"]   & {}                                                  \\
    U_2                      & {\Gamma(U_2, \dtotal)} \arrow[uu, "\iota^*"']
    \end{tikzcd}
\end{equation}
%% sheaf maps over the inclusion rules which is why we care about them ->
the functor \sheaf\ is contravariant because the inclusion map $\iota$ between open sets $U_i \in \mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}. The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$. On the other hand, a set of functions defined on a space $U_2$ must be defined on the subspace $\iota: \Gamma(U_2, \dbase) \rightarrow \Gamma(U_1, \dbase)$ but the reverse does not need to be true. For example, a continuous function defined over the interval $\left[.25,.5\right]$ need not be defined over $\left[0, 1\right]$ but must exist over  $\left[.3, .45\right]$. We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf. We also denote the sheaf over the fiberbundle \dtotal\ as $\sheaf(\dtotal)$. A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}. The sheaf over a limit of open sets that contain a point \dbasepoint\ is approximately the same as a sheaf over a point \dbase\ and is called the stalk $\mathscr{F}_{k}\coloneqq \lim\limits_{U\ni k} \Gamma(U, \dtotal)$ \cite{StalkSheaf2019}. The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$. Our criteria for equivariance is that the set of maps from fiber to fiber $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ is preserved, since these maps include the types of data transformations described in \autoref{sec:related-work:equivariance}. When a fiber bundle is trivial, meaning $\dtotal=\dbase \times \dfiber$, then we can dispense with localization and directly consider the fiber maps $Hom(F_{k}, F_{k})$. In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.

The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$, which we also denote as $\sheaf(\gtotal)$. The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase\, and $\Gamma(\mathcal{S}, \gtotal)$ is the set of sections $\gsection$ over $\mathcal{\gbase}$. The artist is a morphism of sheafs

%is not \Gamma(E), make a category out of it
\begin{equation}
  \label{eq:artist:morphism-of-sheaf}
  \begin{tikzcd}
    & \Gamma(\dtotal) & & & \Gamma(\gtotal) & \\
{} \arrow[rr, "f", Rightarrow, shift right, shorten=3em] & 
& 
{} \arrow[r, "\vartist", shorten=-1.25em, bold] & {} \arrow[rr, "g", Rightarrow, shift right, shorten=3em] & & {} \\
& \mathcal{\dbase}^{op} \arrow[uu, "\mathcal{\dtotal}\coloneqq\sheaf_{\mathcal{\dtotal}}", bend left] 
  \arrow[uu, "\sheaf_{\mathcal{\dtotal}}"', bend right] 
  & & 
& \mathcal{\gbase}^{op} \arrow[uu, "\sheaf_{\mathcal{\gtotal}}", bend left] 
  \arrow[uu, "\mathcal{\gtotal}\coloneqq \sheaf_{\mathcal{\gtotal}}"', bend right] &   
  \end{tikzcd}
\end{equation}

% f is a natural transform, all transforms on the sheafs commute
% g is a natural transform ...
% space of g->g is too big, so we look at Im...
% b/c it's a natural transform, - f carries i over b/c it's a natural transformation

therefore the artist is a map between functors
\begin{equation}
  \tag{\ref{eq:artist:artist}}
  \vartist:\mathcal{\dtotal}\rightarrow{\mathcal{\gtotal}} 
\end{equation} 
which means that the artist is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.

% such that (9)

@@

such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. 
%% such that monoid on this side is monoid on that side

@@

\subsection{Artist Composition}
\label{sec:artist:union}
\begin{figure}[!h]
\centering
\subfloat[+]{\includegraphics[width=.4\columnwidth]{exploding_artist.png}%
\label{fig:artist:union:glue}}
%\hfil
\subfloat[$\oplus$]{\includegraphics[width=.4\columnwidth]{combined_artist.png}%
\label{fig:artist:union:continuity}}
\caption{In \autoref{fig:artist:union:glue}, the outputs of various artist are independently rendered to the same graphic space \gtotal. The inputs to the artists in \autoref{fig:artist:union:continuity} are sections over the same fiber bundle base space \dbase; therefore, they should share transformation functions to produce a consistent image. \note{add a border/make the same size/etc so 1 doesn't run over into two}}
\end{figure}
%% artist composition is fiber products/sums of artists on sheafs -> closed cartesian category maybe?

@@

\section{Construction and formal properties of Artists}
\label{sec:artist:construction}
\begin{figure*}[h!]
  \includegraphics[width=\linewidth]{q.png}
  \caption{}
  \label{fig:constraints:q-overall}
\end{figure*}
